---
title: "Data Sources"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Data Sources}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

We provide access to a variety of different data sources in `conmat`. Most of these are centered around Australian data, as the package was initially created for disease modelling work in Australia. The aim of this vignette is to give a quick tour of the data sources available in `conmat`.

```{r setup}
library(conmat)
```

## Australian Bureau of Statistics (ABS) data

### Accessing Functions

We provide two functions to access (Local Government Area) LGA and state level population age data, which are provided in 5 year age bins from 0, 5, up to 85+

#### `abs_age_lga()`

```{r abs-age-lga}
abs_age_lga(lga_name = "Fairfield (C)")
abs_age_lga(lga_name = "Brisbane (C)")
```

It requires you to know the exact name of the LGA, you can see them in the dataset, `abs_lga_lookup`

```{r abs-lga-lookup}
abs_lga_lookup
```

And if you're not sure about a particular name of a place, you can use `agrep` and `filter` like so:

```{r abs-lga-lookup-filter}
library(dplyr)
abs_lga_lookup %>% 
  filter(agrepl("Sydney", lga))
```


#### `abs_age_state()`

This takes in the abbreviated state names

```{r abs-age-state}
abs_age_state(state_name = "NSW")
```

You can see these state names with:

```{r}
unique(abs_lga_lookup$state)
```

Note that "OT" stands for "other territories"

### ABS data

We provide other ABS data, listed now. You can read the full details of the data at their respective helpfiles, by writing, for example, `?abs_education_state`.

#### Education by state data for 2006 -2020

```{r abs-education-state}
abs_education_state
```

#### Education by state data for 2020

```{r abs-education-state-2020}
abs_education_state_2020
```


#### Employment by LGA for 2016

```{r abs-employ-age-lga}
abs_employ_age_lga
```


#### Number of people in each household by LGA for 2016

```{r abs-household-lga}
abs_household_lga
```


#### LGA age population for 2016 for all states and LGAs

```{r abs-pop-age-lga-2016}
abs_pop_age_lga_2016
```


#### LGA age population for 2020 for all states and LGAs

```{r abs-pop-age-lga-2020}
abs_pop_age_lga_2020
```

#### State age population for 2020

```{r abs-state-age}
abs_state_age
```


## Epidemiology / disease modelling data

### Transmission probabilities from Eyre

A dataset containing data digitised from "The impact of SARS-CoV-2 vaccination on Alpha & Delta variant transmission", by David W Eyre, Donald Taylor, Mark Purver, David Chapman, Tom Fowler, Koen B Pouwels, A Sarah Walker, Tim EA Peto

```{r eyre-transmission-probabilities}
eyre_transmission_probabilities
```

We can visualise the data like so:

```{r eyre-transmission-probabilities-plot}
library(ggplot2)
library(stringr)
library(dplyr)
eyre_transmission_probabilities %>%
   group_by(
     setting,
     case_age_5y,
     contact_age_5y
   ) %>%
   summarise(
     across(
       probability,
       mean
     ),
     .groups = "drop"
   ) %>%
   rename(
     case_age = case_age_5y,
     contact_age = contact_age_5y
   ) %>%
   mutate(
     across(
       ends_with("age"),
       ~ factor(.x,
                levels = str_sort(
                  unique(.x),
                  numeric = TRUE
                )
       )
     )
   ) %>%
   ggplot(
     aes(
       x = case_age,
       y = contact_age,
       fill = probability
     )
   ) +
   facet_wrap(~setting) +
   geom_tile() +
   scale_fill_viridis_c() +
   coord_fixed() +
   theme_minimal() +
   theme(
     axis.text = element_text(angle = 45, hjust = 1)
   )

```
