---
title: "Example Pipeline"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Example Pipeline}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(conmat)
```

This vignette outlines a basic workflow of:

* Create a new synthetic matrix by extrapolating from POLYMOD data to a new age distribution
* Generating a Next Generation Matrix
* Applying Vaccination Rates
* Comparing R0 before and post vaccination rates

## Create a new synthetic matrix from all POLYMOD data

We can create a synthetic matrix from all POLYMOD data by using the `extrapolate_polymod` function. First, let's extract an age distribution from the ABS data.

```{r fairfield}
fairfield_population <- abs_age_lga("Fairfield (C)")
fairfield_population
```

The key features of this age distribution are the columns, `lower.age.limit`, and `population`.

We then extrapolate this using the full POLYMOD data.

```{r extrapolate-fairfield}
synthetic_fairfield_5y <- extrapolate_polymod(
  population = fairfield_population
)
```

By full POLYMOD data, we mean these data:

```{r polymod-setting-population}
polymod_setting <- get_polymod_setting_data()

polymod_population <- get_polymod_population()

polymod_setting
polymod_population
```

The `extrapolate_polymod()` function does the following:

* Fits a model of the contact rate to the full POLYMOD data above
* Predicts it to the provided fairfield population data

It also has options to predict to specified age brackets, defaulting to 5 year age groups up to 75, then 75 and older.

This object, `synthetic_fairfield_5y`, contains a matrix of predictions for each of the settings, home, work, school, other, and all settings.

```{r}
str(synthetic_fairfield_5y)
```

## Generating a Next Generation Matrix

Once infected, a person can transmit an infectious disease to another, creating generations of infected individuals. We can define a matrix describing the number of newly infected individuals in given categories, such as age, for consecutive generations. This matrix is called a "next generation matrix" (NGM).

We can generate an NGM for Australian specific data like so (in a future release we will change this to work for any predicted population)

```{r ngm-fairfield}
ngm_fairfield <- generate_ngm(
  lga_name = "Fairfield (C)",
  age_breaks = c(seq(0, 85, by = 5), Inf),
  R_target = 1.5
)
```

The output of this is a matrix for each of the settings, where each value is the number of newly infected individuals

```{r}
ngm_fairfield$home
str(ngm_fairfield)
```

## Applying Vaccination Rates

Vaccination improves a person's immunity from a disease. When a
sizable section of the population receives vaccinations or when vaccine
coverage is sufficient enough, the likelihood that the unvaccinated
population will contract the disease is decreased. This helps to slow
infectious disease spread as well as lessen its severity. For this reason,
it is important to understand how much of a reduction in probability of
acquisition (the likelihood that an individual will contract the disease),
and probability of transmission (the likelihood that an individual will
spread the disease after contracting it), has occurred as an the effect of
vaccination, in other words the effect of vaccination on the next
generation of infections.

We can use `apply_vaccination()` to return the percentage reduction in acquisition and
transmission in each age group. It does this by taking the outer product
of these reductions in acquisition and transmission by age group, creating
a transmission reduction matrix. The next generation matrices with the
vaccination effects applied are then produced using the obtained
transmission reduction matrix and the next generation matrices passed to
the function as an argument.

```{r}
# Apply vaccination effect to next generation matrices
ngm_nsw_vacc <- apply_vaccination(
 ngm = ngm_fairfield,
 data = vaccination_effect_example_data,
 coverage_col = coverage,
 acquisition_col = acquisition,
 transmission_col = transmission
)
```

## Comparing R0 before and post vaccination rates


##### Example pipeline

Let's predict the average contact rates for ages 0 to Inf in 5 year age breaks in all settings for the NSW LGA, Fairfield

-   Extract the polymod survey population & contact data

```{r}
polymod_contact_data <- get_polymod_setting_data()
polymod_survey_data <- get_polymod_population()
```

-   Fit the contact model to the polymod data for all settings.

```{r}
set.seed(1100)
 setting_models <- fit_setting_contacts(
  contact_data_list = polymod_contact_data,
   population = polymod_survey_data
 )
```

-   Obtain the age wise population of Fairfield.

```{r}
fairfield_age_pop <- abs_age_lga("Fairfield (C)")
fairfield_age_pop

```

-   Obtain the per capita household size of Fairfield

```{r}

fairfield_hh_size <- get_per_capita_household_size(lga = "Fairfield (C)")
fairfield_hh_size
```

-   Extrapolate the fitted model to the Fairfield population data. Predict the contact ages for all combinations of ages in 5 year ages breaks from 0 to 85-Inf.

```{r}

synthetic_settings_5y_fairfield_hh <- predict_setting_contacts(
  population = fairfield_age_pop,
  contact_model = setting_models,
  age_breaks = c(seq(0, 85, by = 5), Inf),
  per_capita_household_size = fairfield_hh_size
)
```

### Transmission Matrices

The initial phase of COVID-19 pandemic saw a significant age dependence in the distribution of confirmed cases with fewer confirmed cases among children. This might have been a result of younger ages being less susceptible to infection and/or less likely to exhibit clinical signs when infected [ [Davies et al](https://www.nature.com/articles/s41591-020-0962-9)] . A population's clinical fraction and age-varying susceptibility to infection both profoundly influence the likelihood that an infectious agent would spread there. This likelihood that an infectious agent would spread from an infected source to a new susceptible host and infect the host is known as transmission probability.

The age and setting specific relative per-contact transmission probability matrices when combined with contact matrices could be used to produce the setting-specific relative next generation matrices (NGMs) which is used to obtain the distribution of numbers of new cases in each generation of infection from any arbitrary initial number of introduced infections. Conmat creates these age and setting specific relative per-contact transmission probability matrices through `get_setting_transmission_matrices()` by making use of clinical fraction and relative susceptibility parameters from Davies et al available through `davies_age_extended`. 
