---
title: "Parallel Computing"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Parallel Computing}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(conmat)
```

`conmat` supports parallelisation via the [`future`](https://future.futureverse.org/) and [`furrr`](https://furrr.futureverse.org/) R packages. The functions that will be impacted by this are:

* extrapolate_polymod
* estimate_setting_contacts
* fit_setting_contacts (called in above functions)
* predict_setting_contacts (called in above functions)

First we set the future plan, saying "multisession", with 4 workers. 

```{r load-future}
library(future)
plan(multisession, workers = 4)
```

Then we run our code as normal to get the parallelisation! (note that you must specify the plan, otherwise it does not know how to parallelise. See the [future package documentation](https://future.futureverse.org/reference/plan.html) for more details).

Note that these functions will run about 3 times faster than normal, they might still take some time. They are able to run in parallel as we are fitting a model to each setting, which is a task that is embarrasingly parallel.

```{r show-off-furrr-1}
perth <- abs_age_lga("Perth (C)")
perth_contacts <- extrapolate_polymod(
  population = perth
)
```

```{r}
settings_estimated_contacts <- estimate_setting_contacts(
  contact_data_list = get_polymod_setting_data(),
  survey_population = get_polymod_population(),
  prediction_population = get_polymod_population(),
  age_breaks = c(seq(0, 75, by = 5), Inf),
  per_capita_household_size = NULL
)
```

```{r}
polymod_setting_data <- get_polymod_setting_data()
polymod_population <- get_polymod_population()

contact_model <- fit_setting_contacts(
  contact_data_list = polymod_setting_data,
  population = polymod_population
)
```

```{r}
synthetic_settings_5y_perth <- predict_setting_contacts(
  population = perth,
  contact_model = contact_model,
  age_breaks = c(seq(0, 85, by = 5), Inf)
)
```

For computational purposes, we need to ensure we shut down the parallel workers, as per the vignette from `future`, [Best Practices for Package Developers](https://future.futureverse.org/articles/future-7-for-package-developers.html#writing-examples):

```{r}
## Shut down parallel workers
future::plan("sequential")
```

