---
title: 'conmat: generate synthetic contact matrices for a given age-stratified population'
authors:
- affiliation: 1
  name: Nicholas Tierney
  orcid: 0000-0003-1460-8722
- affiliation: 1,2
  name: Nick Golding
  orcid: 0000-0001-8916-5570
- affiliation: 1,3
  name: Aarathy Babu
  orcid: 
- affiliation: 4
  name: Michael Lydeamore
  orcid: 0000-0001-6515-827X
- affiliation: 1,3
  name: Chitra Saraswati
  orcid: 0000-0002-8159-0414
date: "03 May 2024"
output:
  html_document:
    keep_md: yes
  pdf_document: default
bibliography: references.bib
tags:
- epidemiology
- R
- infectious disease
affiliations:
- index: 1
  name: Telethon Kids Institute
- index: 2
  name: Curtin University
- index: 3
  name: 
- index: 4
  name: Monash University
---

```{r setup}
#| label: setup
#| echo: false
#| message: false
#| warning: false
#| comment: false
knitr::opts_chunk$set(comment = "#>",
                      echo = TRUE,
                      out.width = "95%",
                      retina = 3,
                      fig.height = 4,
                      fig.align = "center",
                      dpi = 300,
                      dev = "png")
options(tinytex.clean = FALSE)
```

[comment]: Revised, edited text below the original text.  

# Summary

#TODO - A better first sentence that encapsulates conmat use? 
Understanding the dynamics of infectious disease transmission in a population is an important task (?) for epidemiologists and public policy makers.
Identifying vulnerable groups and predicting disease transmission (?)dynamics are essential for informed public health decision-making. Infectious diseases such as influenza and coronavirus spread through human-to-human interactions, or in other words, "social contact". Quantifying social contact and its patterns can provide critical insights into how diseases spread.

We can measure social contact through social contact surveys, where people describe the number and type of social contact they have. These surveys provide (?) a measure of contact rates: an empirical estimate of the number of social contacts from one age group to another and the setting of contact. For example, we might learn from a contact survey that homes have higher contact between 25-50 year olds and 0-15 year olds, whereas workplaces might have high contact within 25-60 year olds.

These social contact surveys exist for a few countries. As an example, the "POLYMOD" study by @mossong2008 covered 8 European countries: Belgium, Germany, Finland, Great Britain, Italy, Luxembourg, The Netherlands, and Poland [@mossong2008]. However, what do we do when we want to (?)look at/estimate contact rates in other countries that have not been measured? We can use this existing data to help us project to countries or places that do not have empirical contact survey data. These are called "synthetic contact matrices". A popular approach by @prem2017 projected contact rates from the POLYMOD study to 152 countries. This was later updated to include synthetic contact matrices for 177 countries at "urban" and "rural" levels for each country [@prem2021]. 

However, there were major limitations with the methods in @prem2021. First, not all countries were included in their analyses. Second, some of the synthetic contact matrices did not have enough granularity; they covered a large area, such as "urban" or "rural" for a given country. This is disadvantegous as public health groups might need to make predictions for more fine-grained areas within a country, such as a district or municipality. Third, the methodology used by Prem et al. was challenging to reuse in other contexts. Prem et al. provided the code used for their analysis, but that code was not designed for replicability and user-defined inputs.

The `conmat` package was created to fill a specific need for creating synthetic contact matrices for specific local government areas for Australia, for work commissioned by the Australian government. We created methods and software to facilitate the following:

The `conmat` package was developed to fill the specific need of creating synthetic contact matrices for local government areas in Australia. This package is used for [ #TODO what work, specifically? Health? Provide example. Or is *this* package commissioned by the Aus govt? ] work commissioned by the Australian government.
We developed methods and software to facilitate the following tasks.

- Generate, as output, synthetic contact matrices from age-stratfied population data.
- Create next generation matrices (NGMs).
- Apply vaccination reduction to NGMs.
- Use NGMs in disease modelling.
- Provide tidied population data from the Australian Bureau of Statistics.

# Example

As an example, let us generate a contact matrix for a local government area within Australia, using a model fitted from the POLYMOD data.

Suppose we want to generate a contact matrix for the City of Perth. We can get the age-stratified population data for Perth from the helper function `abs_age_lga`:

```{r}
#| label: load-conmat
library(conmat)
perth <- abs_age_lga("Perth (C)")
perth
```

We can generate a contact matrix for `perth` using the `extrapolate_polymod` function, where the contact matrix is generated using a model fitted from the POLYMOD data. 

```{r}
#| label: extrapolate-polymod
#| echo: true
perth_contact <- extrapolate_polymod(population = perth)
perth_contact
```

We can plot the resulting contact matrix for Perth with `autoplot`:

```{r}
#| label: autoplot-contacts
autoplot(perth_contact)
```

# Implementation

`conmat` was built to predict at four settings: work, school, home, and other. 
The model fitted is a Poisson generalised additive model (GAM) which predicts the count of contacts, with an offset for the log of participants. The model has six (?)covariates/terms to explain six key features of the relationship between ages, and two optional terms for attendance at school or work. The two optional terms are included depending on which setting the model is fitted for. There are four resulting models, with one model fitted for each setting. 

The six terms are the `|i-j|`, 

The six key features of the relationship between the age groups, represented by the six terms, are displayed in the figure below.

```{r}
# use DHARMA to show a partial dep plot of the six main terms
```

Each cell in the resulting contact matrix, indexed *i*, *j*, is the predicted number of people in age group *j* that a single individual in age group *i* will have contact with per day. If you sum across all the *j* age groups for each *i* age group, you get the predicted total number of contacts per day for each individual of age group *i*. [ #TODO expected, predicted, or average? ]

[ #TODO notes-to-self: the model structure wasn't generated through any particularly robust process, it was just coming up with structures that looked mildly appropriate for our use case. ]

The results are surveys that can be found in the syncomat package, which can be seen at: https://idem-lab

## Model interfaces

We provide functions for model fitting at various use cases. Further detail for each of the following functions can be seen at: https://idem-lab.github.io/conmat/dev/

* `fit_single_contact_model()`
    * Fits a generalised additive model (GAM) using contact survey data and population size information. This function is recommended for when you want to fit a model to only one setting, for which you might want to provide your own contact survey data.

* `predict_contacts()`

    * This takes a fitted model from `fit_single_contact_model()` and predicts [ #TODO what is predicted? ] to a provided (?) population structure.

* `fit_setting_contacts()`
    * Fits the `fit_single_contact_model()` to each setting. This function is useful for when you have multiple settings to fit. Returns a list of fitted models. 

* `predict_setting_contacts()`
    * Takes a list of fitted models from `fit_setting_contacts()` and predicts [ #TODO what is predicted? ] to a given population for each setting.

* `estimate_setting_contacts()`
    * A convenience function that fits multiple models, one for each setting. This means fitting `fit_setting_contacts()` and then `predict_setting_contacts()`. Recommended for when you have multiple settings to fit and want to predict to a given population as well.

* `extrapolate_polymod()`
    * Takes population information and projects pre-fit model from POLYMOD - used for speed when you know you want to take an already fit model from POLYMOD and just fit it to your provided population.

[ #TODO for the above it's good to explain what exactly is predicted. Otherwise it's confusing for the user to understand what each of the model outputs? ]

# Conclusions and Future Direction

* Create a contact matrix using a custom contact survey from another source, such as the `socialmixr` R package.
* Predict to any age brackets - such as monthly ages, for example, 1, 3, 6, month year old infants
* Add ability to fit multiple contact surveys at once, e.g., POLYMOD and another dataset
* Add ability to include known household age distributions as offsets in the 'home' setting model, in place of the whole population distribution. So compute household age matrices (like age-structured contact matrices, but for household members instead of contacts) from POLYMOD data. If we compute a different one for each household size, in the POLYMOD data (probably estimated with another GAM, to make best use of the limited data) we might be able to extrapolate household age matrices to new countries based on the distribution of household sizes.
* Add methods for including household size distributions
* Add uncertainty to estimates
* Move Australian centric data into its own package
* Add documentation on specifying your own GAM model and using this workflow

# References
